Biogeography \& Environmental Conditions Shape Phage \& Bacteria Interaction Networks Across the Human Microbiome

### 2017-08-11 revfiles/ReviewerResponse.md

Taken together, the machine leaning-based approach would only require some additional tests to exclude or estimate the degree of bacterial contamination in the VLP fraction and some explanations on how abundances in VLP and microbial fractions were combined and normalized. With regard to the application of the methodology, two of the three datasets are severely underpowered to support major claims on the impact of diet and health state. I would thus recommend to put the focus of the manuscript on the analysis of the well-powered skin data set and reduce the diet and obesity-related part as potentially promising case-examples while clearly describing the limitations of the analysis (which is done) and toning down or removing some of the respective claims.

## Major Points

1.  Conceptual distinction between infectiveness and community stability: For the development of the network approach, the authors use infectiveness data as evidence for interactions between phage and bacteria. For the analysis part, such interactions are used to describe community stability (i.e., more interactions per node, higher stability). If phages are considered as agents grazing on their microbial hosts, is it justified to use degree of infectiveness analogous to community stability? Some discussion or background information would be helpful.

*Thank you for raising this point which warrants further discussion. We clarified in the text why we feel this is justified, as the degree of stability refers to the reduced likelihood that a randomly removed node will cause a major divide or lack of connection between agents within the network. The infectiveness refers to the link we drew between the bacteria and phages (how we connected phages and bacteria in our networks). We also provided citations to work that further explains this in metabolic and phage-bacteria systems, if the readers are interested in a more in-depth discussion of the topic.* **LINES 180 - 187; 268 - 270**

@@

2.  Testing for bacterial DNA in the phage fraction: The authors do not describe how bacterial sequences in the VLP fraction were eliminated. If bacterial sequences were present, they would be labeled as phages and lead to false predictions.

*We agree that this needed to be better outlined in the text. It is important that we make clear that we are not violating our assumption that that phage OGU pool represents phages and the bacterial OGU pool represented bacteria. We updated the manuscript to address this concern in multiple ways. First, because this was a post hoc analysis of published data, we do not have access to the samples for performing further molecular QC. The original study authors did however perform and report a variety of QC measures to confirm the purity of their virome sample sets. To this end, we summarized those QC results into Table S1, which we added to the manuscript and referenced in the results. Furthermore, we performed our own local alignment (blast) QC protocol to assess the degree to which our OGUs were contaminated. We found that many phages had genomic elements similar to reference bacteria genomes, which agrees with the previous findings that the majority of the phages in these systems are temperate. We also identified two complete phages using the stringent Virsorter algorithm. Conversely, there were minimal phage reference genome elements identified in the bacterial OGUs, which also agrees with the previous findings that these sequences are mostly bacterial (cited in Table S1). We also did not identify any phages in the bacterial dataset using Virsorter. The lack of prophage identification, both here and in the previous study analyses, speaks to the need to sequence the bacterial fraction more deeply and create more robust assemblies (their larger size prevents assembly completion at the level that can be done with viruses).* **TABLE S1; LINES 92 - 121; 379 - 390**

@@

3.  Combining data from VLPs and microbial fractions: Information on how data from these two individual samples were matched and normalized is not provided. How are potentially different amounts of VLPs in different samples accounted for?

*We are unclear what the reviewer was asking. Our goal was to use the edge weights to reflect samples that were highly relative abundant in both the bacterial metagenome and the phage metagenome. The edge weight used was therefore a multiplication function of the two relative abundance values, yielding a high value for those that were highly abundant in both, and low values for those that were not. This is described in the methods section of the manuscript. These values were recorded on a per-sample basis (within the subgraphs) so that each potential combination is accounted for, within each sampling pair. Furthermore, it may be worth noting that the bacterial and viral DNA extractions were done separately so as to maximize purity of the samples. This approach allows us to more accurately measure the relative abundance in the bacterial and especially viral samples, as has been noted previously. We clarified this in the text and added more relevant citations.* **LINES 151 - 155; 157 - 159**

@@

4.  Datasets underpowered to sustain claims on diet and obesity: Although the authors point out that the data are underpowered, they still make strong claims and the analysis of these data as major points of this manuscript.

For example, 
a.  the authors describe (158-164) how diet affects the network structure, without using sufficient data points to perform any statistical tests.
b.  the authors claim (166-169) a difference between obese and healthy individuals by using one or two data points per group.
With respect to these analyses, the authors write for example, "We found evidence that diet was sufficient to alter gut microbiome network connectivity" (235), which is a clear overstatement and such claims should be modified here and elsewhere in the manuscript.

*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short. Our goal was to present these as interesting observations, somewhat like a case study. We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29; 185 - 206; 274 - 278**

@@

## Minor Points

* Throughout the manuscript, it is hard to follow how which and how many data points were used for each analysis. A supplementary table listing the samples for each of the studies and providing sample sizes in each figure would be very helpful.

*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

@@

* Figure 2 could use headers above the plots to indicate which one relate to impact of diet (AB) and which one to obesity (CD).

*This was especially confusing because of the mislabeled axes. We fixed the axis labels to make the data clearer.* **FIGURE 2**

* Figure 2 panels C and D have wrong x-axis labels. Should be "obese" and "healthy". Here it is not clear why only 2 and 1 data points are available. The original data in Reyes et al. seem to consist of 32 fecal samples from 13 individuals from 5 families with at least two time points.

*We fixed the error, and made the causes for the sample size clearer in the text as well as the figure legends. Thank you for pointing this out.* **LINES 200 - 202; FIGURE LEGEND 2**

@@

* Figure 3A. Not clear why only 8 data points are available if Minot et al. report 15 samples from 5 individuals.

*We only used the subset of data points that had temporal data spaced approximately the same distance apart. We made this clearer in the text.* **LINES 191 - 192; 224 - 225; FIGURE LEGEND 1,2**

@@

* Figure 3B. Not clear how data were computed. If the comparison of within individuals (i.e. 1 value if same data as in 3A) with mean of all other individuals, then standard deviations should be shown. Suggest to show only the smallest inter-individual distance, which would greatly strengthen the point on individuality.

*Yes, this is a comparison between the interpersonal distance, and the distance between the person's samples and all other samples. We considered calculating the distance differences using standard deviation and pairing the samples as every interpersonal comparison instead of the average. The problem we found was that that pairwise comparison inflated the significance values by artificially creating a large n, all of which were not truly independent. We therefore collapsed the data points into a single average instead of pairing the same sample to many other samples. Showing the smallest distance is also an interesting idea, but we are worried that this may also skew the data by failing to account for the variation in the sample distributions.*

* Figure 3C. Not clear which data were used, in particular if same data as in Figure 2 were used. E.g., were only the twins included, or also the mothers?

*We made this information clearer in the text.* **LINES 229 - 230; FIGURE LEGEND 3**

@@

## Major comments:

1. Network inference: operational genomic units were linked sample-wise, microbiome type wise or all microbiome together? Aminoacid similarity is not specific enough and is likely the cause of the high density of the network, which suggests that almost all phages infect all bacteria (86% density), meaning the resulting network is (likely) no statistically different from a random network with equal number of links. In the absence of further evidence and controls, I am not convinced with this result.

*We would like to thank the reviewer for pointing out that this area needed clarification. The operational genomic units were linked sample-wise, by linking the OGUs found between each sample. The source of these samples were recorded, so that we knew whether they originated from the gut or the skin, etc. We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model. As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest. It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies. In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 103 - 108; 311 - 329**

@@

2. Number of samples per study and metadata groups are not reported but the authors state about the diet effect: "Tests for statistical differences were not performed due to the small sample size ". If sample sizes are small for statistical analysis, no comparison should be made in the first place. The plots of Figure 2 feature 2 vs. 3 comparison for high fat vs. low fat and 2 vs. 1 for obese vs. healthy...are these the nb. of samples per group? Bacterial or phage diversity are likely confounders for network connectivity comparisons and this is overlooked here.

*We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions. We also needed to clarify the analysis being performed. It is also true that the network analysis is a function of the diversity of the two communities, and we did not wish to make it sound like they are separate. We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community. We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

@@

1.  Training set contained on 43 diverse bacterial species and 30 diverse phage strains. Given than these numbers are (much) lower than the ones published in references 41, 47-51, from which the data on known interactions was extracted, I wonder how the curation/filtering was done.

*Our limitation was not on how many known interactions that could be found, but instead on how many non-interactions could be found. In other words, we needed to utilize known infectious pairs, as well as phages that have been validated as not infecting certain bacteria. Because we were limited to only a small number of validated non-interactions, we decided to also limit the number of positive interactions included, so as to prevent a large bias in the ratio between the two interaction types in the training dataset. We clarified that text.* **LINES 127 - 129**

2.  The sensitivity of the classifier is reported as 0.846. However, "Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction". This means the sensitivity can not be higher than 0.67...

*Here we are reporting the accuracy for those values that were scored. Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored. We have clarified this point.* **SEE LINES 145 - 146; FIGURE LEGEND 1**

@@

3.  Dataset: The authors should report the number of samples per study and associated metadata (e.g low /high fat diet, obese/lean).

*We have added this information.* **LINES 191 - 192; 200 - 202; 219 - 221; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

@@

5.  Network construction: The master network contained the three studies as sub-networks, which themselves each contained sub-networks for each sample. The authors provide only numbers of phage and bacterial nodes for the master network. For predicting the associations, were all phages and bacteria considered disregarding the sample type (skin vs. gut)?

*Yes, this is described correctly by the reviewer.*

6.  The phages and bacteria in the gut diet and twin sample sets were more sparsely related: each contained fewer than 150 vertices, fewer than 20,000 relationships". This is a bipartite network, what should be reported is the number of vertices of each type rather than the total nb of vertices (this is relevant to calculate the network density, we cannot know here what is the nb of possible interactions).

*We added these values to the results section of the manuscript so that the reader can access this information.* **LINES 155 - 165**

@@

7.  Comparison with Paez-Espino et al. (Nature 2016, doi:10.1038/nature19094) phage-host human associated subnetworks should be done.

*This is an intriguing suggestion. It sounds like this comment is a reference to the host range inferences within the suggested manuscript. Paez-Espino et al used CRISPR and tRNA comparisons to identify bacterial hosts for the bacteriophage genomes and genome fragments. The group's work suggested that phage host range is broader than has been traditionally appreciated. In our discussion section, we mentioned a comparison of these and related findings to our own dataset. A reference to this information was included in the operational genomic unit section of the discussion.* **LINES 318**

@@

The question is how were the above biases (including the kitome) taken into consideration while conducting these analyses. The authors may be better off carrying out the analysis only using data from references 13 and 14 for skin and only data from 38 for gut?

*There certainly could have been extraction, kit, or other related differences between the 38 and 39 dataset. We feel these findings, which we attempt to present more as case observations that will require more dedicated study, are worth noting in the paper. We altered the text to better clarify our position. We also avoided analyses in which we directly compared datasets, because we are unable to discern between true biological differences and differences between the studies. Such analyses would require further and properly controlled sampling.*

Even though the authors have acknowledged some of these biases including MDA, in their discussion, these concerns should be explicitly discussed and measures they have taken to overcome these.

*We included a description of these biases associated with these methods. As far as measures to overcome these biases, we acknowledge the biases and only performed analyses within studies where samples were treated the same. The goal is for this work to use these findings as a jumping off point, and performing future studies with more limited biases from techniques such as MDA.* **SEE LINES 302 - 310**

### 2017-08-11 revfiles/ReviewerResponse.md

*We appreciate this point referring to figure 2, and while we attempted to avoid too strong of claims, we realize we fell short. Our goal was to present these as interesting observations, somewhat like a case study. We addressed this in the manuscript by further minimizing the effects of our claims and adding relevant caveats.* **LINES 29; 185 - 206; 274 - 278**

## Minor Points

* Throughout the manuscript, it is hard to follow how which and how many data points were used for each analysis. A supplementary table listing the samples for each of the studies and providing sample sizes in each figure would be very helpful.

*We added the relevant sample size information to the text and figure legends to make clearer what the data actually represent.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

@@

*We would like to thank the reviewer for pointing out that this area needed clarification. The operational genomic units were linked sample-wise, by linking the OGUs found between each sample. The source of these samples were recorded, so that we knew whether they originated from the gut or the skin, etc. We also agree that amino acid similarity of genes along is not enough, and that is why we incorporated other similarity metrics in our random forest model. As we described, even though the amino acid similarity metric was the most "important" to the model, it was used within the context of the other metrics in the random forest. It is true that this is a dense network, and we hypothesize that that may be because each OGU does not necessarily represent a specific phage or bacterial strain, but is an operationally defined group of similar genomes, similar to the operation taxonomic unit (OTU) often used in microbiome studies. In other words, we are not using the OGUs as individual phage strains, but instead these OGUs are operational units that allow us to condense our dataset by contig similarity.* **SEE LINES 103 - 108; 311 - 329**

2. Number of samples per study and metadata groups are not reported but the authors state about the diet effect: "Tests for statistical differences were not performed due to the small sample size ". If sample sizes are small for statistical analysis, no comparison should be made in the first place. The plots of Figure 2 feature 2 vs. 3 comparison for high fat vs. low fat and 2 vs. 1 for obese vs. healthy...are these the nb. of samples per group? Bacterial or phage diversity are likely confounders for network connectivity comparisons and this is overlooked here.

*We appreciate this point and while we attempted to report this as more of a case trial observation (as outlined for the related comment above), we unintentionally used too strong of language, thus overstating the finding beyond our intentions. We also needed to clarify the analysis being performed. It is also true that the network analysis is a function of the diversity of the two communities, and we did not wish to make it sound like they are separate. We would be hesitant however to call this dependence a confounder since it is more building off of the diversity data and incorporating other additional information from the community. We made this clearer in the text.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

@@

2.  The sensitivity of the classifier is reported as 0.846. However, "Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction". This means the sensitivity can not be higher than 0.67...

*Here we are reporting the accuracy for those values that were scored. Pairs without scores were omitted from this analysis, and we instead only focused on those that were scored. We have clarified this point.* **SEE LINES 145 - 146; FIGURE LEGEND 1**

3.  Dataset: The authors should report the number of samples per study and associated metadata (e.g low /high fat diet, obese/lean).

*We have added this information.* **LINES 191 - 192; 200 - 202; 224 - 225; 229; FIGURE LEGENDS 2,3,4**

### 2017-08-29 doc/manuscript.Rmd
A couple minor updates.
Updates for resubmission.

```{r loadfiles, echo=FALSE}
sigfig = 3
genfigurestats <- read.delim(file = "../rtables/genfigurestats.tsv", header = TRUE, sep = "\t")
nestats <- read.delim(file = "../rtables/nestats.tsv", header = TRUE, sep = "\t")
sitestats <- read.delim(file = "../rtables/sitestattable.tsv", header = TRUE, sep = "\t")
genmodelper <- read.delim(file = "../rtables/genmodelper.tsv", header = TRUE, sep = "\t")
interstats <- read.delim(file = "../rtables/interstats.tsv", header = TRUE, sep = "\t")
skinboxsig <- read.delim(file = "../rtables/skinboxsig.tsv", header = TRUE, sep = "\t")
```

@@

Four phage and bacterial genomic features were used in a random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. The resulting random forest model was assessed and the area under its receiver operating characteristic (ROC) curve was `r signif(genmodelper$ROC, digits = sigfig)`, the model sensitivity was `r signif(genmodelper$Sens, digits = sigfig)`, and specificity was `r signif(genmodelper$Spec, digits = sigfig)` **(Figure \ref{RocCurve} A)**. The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the minimal amount of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

The four scores were used as parameters in a random forest model to classify bacteria and bacteriophage pairs as either having infectious interactions or not. The classification model was built using the Caret R package (v6.0.73) [@caretClassificatio:_U2Lit_1]. The model was trained using five-fold cross validation with ten repeats. Pairs without scores were classified as not interacting. The model was optimized using the ROC value. The resulting model performance was plotted using the plotROC R package.

@@

![**Summary of Multi-Study Network Model.** *(A) Average ROC curve used to create the microbiome-virome infection prediction model. (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages. The importance score is defined as the mean decrease in accuracy of the model when a feature (e.g. Pfam) is excluded. Features include the local gene alignments between bacteria and phage genes (denoted `blastx`; the blastx algorithm in Diamond aligner), local genome nucleotide alignments between bacteria and phage OGUs, presence of experimentally validated protein family domains (Pfams) between phage and bacteria OGUs, and CRISPR targeting of bacteria toward phages (CRISPR). (C) Proportions of samples included (gray) and excluded (red) in the model. Samples were excluded from the model because they did not yield any scores. Those interactions without scores were automatically classified as not having interactions. (D) Bipartite visualization of the resulting phage-bacteria network. Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines. This network includes information from all three published studies. (E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }

@@

Table: Summary of the primary quality control measures reported in the original publications of the viromes used in this current study.

>>>>

```{r loadfiles, echo=FALSE}
sigfig = 3
genfigurestats <- read.delim(file = "../rtables/genfigurestats.tsv", header = TRUE, sep = "\t")
nestats <- read.delim(file = "../rtables/nestats.tsv", header = TRUE, sep = "\t")
sitestats <- read.delim(file = "../rtables/sitestattable.tsv", header = TRUE, sep = "\t")
genmodelper <- read.delim(file = "../rtables/genmodelper.tsv", header = TRUE, sep = "\t")
interstats <- read.delim(file = "../rtables/interstats.tsv", header = TRUE, sep = "\t")
skinboxsig <- read.delim(file = "../rtables/skinboxsig.tsv", header = TRUE, sep = "\t")
averageauc <- read.delim(file = "../data/avgaucnested.tsv", header = TRUE, sep = "\t")
```

@@

Four phage and bacterial genomic features were used in a random forest model to predict infectious relationships between bacteria and phages: 1) genome nucleotide similarities, 2) gene amino acid sequence similarities, 3) bacterial Clustered Regularly Interspaced Short Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4) similarity of protein families associated with experimentally identified protein-protein interactions [@Orchard:2014hq]. The resulting random forest model was assessed using nested cross validation, and the median area under its receiver operating characteristic (ROC) curve was `r signif(averageauc[c(averageauc$type %in% "AUC"), 2], digits = sigfig)`, the median model sensitivity was `r signif(averageauc[c(averageauc$type %in% "SENS"), 2], digits = sigfig)`, and median specificity was `r signif(averageauc[c(averageauc$type %in% "SPEC"), 2], digits = sigfig)` **(Figure \ref{RocCurve} A)**. The most important predictor in the model was amino acid similarity between genes, followed by nucleotide similarity of whole genomes **(Figure \ref{RocCurve} B)**. Protein family interactions were moderately important to the model, and CRISPRs were largely uninformative, due to the minimal amount of identifiable CRISPRs in the dataset and their redundancy with the nucleotide similarity methods **(Figure \ref{RocCurve} B)**. Approximately one third of the training set relationships yielded no score and therefore were unable to be assigned an interaction prediction **(Figure \ref{RocCurve} C)**.

@@

The four scores were used as parameters in a random forest model to classify bacteria and bacteriophage pairs as either having infectious interactions or not. The classification model was built using the Caret R package (v6.0.73) [@caretClassificatio:_U2Lit_1]. The model was trained using five-fold cross validation with ten repeats, and the median model performance was evaluated by training the model on 80% of the dataset and testing performance on the remaining 20%. Pairs without scores were classified as not interacting. The model was optimized using the ROC value. The resulting model performance was plotted using the plotROC R package.

@@

![**Summary of Multi-Study Network Model.** *(A) Median ROC curve (dark red) used to create the microbiome-virome infection prediction model, based on nested cross validation over 25 random iterations. The maximum and minimum performance are shown in light red. (B) Importance scores associated with the metrics used in the random forest model to predict relationships between bacteria and phages. The importance score is defined as the mean decrease in accuracy of the model when a feature (e.g. Pfam) is excluded. Features include the local gene alignments between bacteria and phage genes (denoted `blastx`; the blastx algorithm in Diamond aligner), local genome nucleotide alignments between bacteria and phage OGUs, presence of experimentally validated protein family domains (Pfams) between phage and bacteria OGUs, and CRISPR targeting of bacteria toward phages (CRISPR). (C) Proportions of samples included (gray) and excluded (red) in the model. Samples were excluded from the model because they did not yield any scores. Those interactions without scores were automatically classified as not having interactions. (D) Bipartite visualization of the resulting phage-bacteria network. Phage OGUs are presented in orange, bacteria OGUs in red, and their interaction edges are represented as connecting lines. This network includes information from all three published studies. (E) Network diameter (measure of graph size; the greatest number of traversed vertices required between two vertices), (F) number of vertices, and (G) number of edges (relationships) for the total network (orange) and the individual study sub-networks (diet study = red, skin study = yellow, twin study = green).* \label{RocCurve}](){ width=85% }

@@

Table S1: Summary of the primary quality control measures reported in the original publications of the viromes used in this current study.

### 2017-08-29 doc/manuscript.tex
A couple minor updates.
Updates for resubmission.

Four phage and bacterial genomic features were used in a random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions (53). The resulting random forest model was
assessed and the area under its receiver operating characteristic (ROC)
curve was 0.846, the model sensitivity was 0.829, and specificity was
0.767 \textbf{(Figure \ref{RocCurve} A)}. The most important predictor
in the model was amino acid similarity between genes, followed by
nucleotide similarity of whole genomes \textbf{(Figure \ref{RocCurve}
B)}. Protein family interactions were moderately important to the model,
and CRISPRs were largely uninformative, due to the minimal amount of
identifiable CRISPRs in the dataset and their redundancy with the
nucleotide similarity methods \textbf{(Figure \ref{RocCurve} B)}.
Approximately one third of the training set relationships yielded no
score and therefore were unable to be assigned an interaction prediction
\textbf{(Figure \ref{RocCurve} C)}.

@@

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) (79). The model was trained using five-fold cross
validation with ten repeats. Pairs without scores were classified as not
interacting. The model was optimized using the ROC value. The resulting
model performance was plotted using the plotROC R package.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.85000\textwidth]{}
\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A)
Average ROC curve used to create the microbiome-virome infection
prediction model. (B) Importance scores associated with the metrics used
in the random forest model to predict relationships between bacteria and
phages. The importance score is defined as the mean decrease in accuracy
of the model when a feature (e.g.~Pfam) is excluded. Features include
the local gene alignments between bacteria and phage genes (denoted
\texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
nucleotide alignments between bacteria and phage OGUs, presence of
experimentally validated protein family domains (Pfams) between phage
and bacteria OGUs, and CRISPR targeting of bacteria toward phages
(CRISPR). (C) Proportions of samples included (gray) and excluded (red)
in the model. Samples were excluded from the model because they did not
yield any scores. Those interactions without scores were automatically
classified as not having interactions. (D) Bipartite visualization of
the resulting phage-bacteria network. Phage OGUs are presented in
orange, bacteria OGUs in red, and their interaction edges are
represented as connecting lines. This network includes information from
all three published studies. (E) Network diameter (measure of graph
size; the greatest number of traversed vertices required between two
vertices), (F) number of vertices, and (G) number of edges
(relationships) for the total network (orange) and the individual study
sub-networks (diet study = red, skin study = yellow, twin study =
green).} \label{RocCurve}}
\end{figure}

>>>>

Four phage and bacterial genomic features were used in a random forest
model to predict infectious relationships between bacteria and phages:
1) genome nucleotide similarities, 2) gene amino acid sequence
similarities, 3) bacterial Clustered Regularly Interspaced Short
Palindromic Repeat (CRISPR) spacer sequences that target phages, and 4)
similarity of protein families associated with experimentally identified
protein-protein interactions (53). The resulting random forest model was
assessed using nested cross validation, and the median area under its
receiver operating characteristic (ROC) curve was 0.788, the median
model sensitivity was 0.952, and median specificity was 0.615
\textbf{(Figure \ref{RocCurve} A)}. The most important predictor in the
model was amino acid similarity between genes, followed by nucleotide
similarity of whole genomes \textbf{(Figure \ref{RocCurve} B)}. Protein
family interactions were moderately important to the model, and CRISPRs
were largely uninformative, due to the minimal amount of identifiable
CRISPRs in the dataset and their redundancy with the nucleotide
similarity methods \textbf{(Figure \ref{RocCurve} B)}. Approximately one
third of the training set relationships yielded no score and therefore
were unable to be assigned an interaction prediction \textbf{(Figure
\ref{RocCurve} C)}.

@@

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) (79). The model was trained using five-fold cross
validation with ten repeats, and the median model performance was
evaluated by training the model on 80\% of the dataset and testing
performance on the remaining 20\%. Pairs without scores were classified
as not interacting. The model was optimized using the ROC value. The
resulting model performance was plotted using the plotROC R package.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.85000\textwidth]{}
\caption{\textbf{Summary of Multi-Study Network Model.} \emph{(A) Median
ROC curve (dark red) used to create the microbiome-virome infection
prediction model, based on nested cross validation over 25 random
iterations. The maximum and minimum performance are shown in light red.
(B) Importance scores associated with the metrics used in the random
forest model to predict relationships between bacteria and phages. The
importance score is defined as the mean decrease in accuracy of the
model when a feature (e.g.~Pfam) is excluded. Features include the local
gene alignments between bacteria and phage genes (denoted
\texttt{blastx}; the blastx algorithm in Diamond aligner), local genome
nucleotide alignments between bacteria and phage OGUs, presence of
experimentally validated protein family domains (Pfams) between phage
and bacteria OGUs, and CRISPR targeting of bacteria toward phages
(CRISPR). (C) Proportions of samples included (gray) and excluded (red)
in the model. Samples were excluded from the model because they did not
yield any scores. Those interactions without scores were automatically
classified as not having interactions. (D) Bipartite visualization of
the resulting phage-bacteria network. Phage OGUs are presented in
orange, bacteria OGUs in red, and their interaction edges are
represented as connecting lines. This network includes information from
all three published studies. (E) Network diameter (measure of graph
size; the greatest number of traversed vertices required between two
vertices), (F) number of vertices, and (G) number of edges
(relationships) for the total network (orange) and the individual study
sub-networks (diet study = red, skin study = yellow, twin study =
green).} \label{RocCurve}}
\end{figure}
